---
title: "Regression Models Course Project"
author: "Eduardo Alvela Jr."
date: "October 15, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:
"Is an automatic or manual transmission better for MPG"
"Quantify the MPG difference between automatic and manual transmissions"


Load the Data

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).
A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg	Miles/(US) gallon

[, 2]	cyl	Number of cylinders

[, 3]	disp	Displacement (cu.in.)

[, 4]	hp	Gross horsepower

[, 5]	drat	Rear axle ratio

[, 6]	wt	Weight (1000 lbs)

[, 7]	qsec	1/4 mile time

[, 8]	vs	Engine (0 = V-shaped, 1 = straight)

[, 9]	am	Transmission (0 = automatic, 1 = manual)

[,10]	gear	Number of forward gears

[,11]	carb	Number of carburetors



```{r mtcars}
library(datasets)
data(mtcars)
head(mtcars)
```

# Exploratory Analysis

Testing a hypothesis if automatic and manual transmission have the same average in MPG
Using t-test

```{r mtcars2}
result <- t.test(mtcars$mpg ~ mtcars$am)
result
```

```{r mtcars3}
result$p.value
```

Since the p-value is 0.00137, we reject out null hypothesis. Automatic and manual transmissions are from different populations.

Difference:

```{r mtcars4}
result$estimate
```

```{r mtcars5}
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
boxplot(mpg ~ am,
data=mtcars,
ylab="Miles Per Gallon (MPG)",
xlab="Transmission Type 0 (Auto) 1 (Manual)",
col=(c("red","blue")))
```

This graph shows that there is a significant increase in MPG for vehicles with a manual vs automatic.

```{r mtcars6}
fit <- lm(mpg ~ factor(am), data=mtcars)
summary(fit)
```

The adjusted R-squared value is around 34% of the regression variance which explained by the model. 
Let show how other predictor variables will
impact:

```{r mtcars7}
data(mtcars)
fit2 <- lm(mpg ~ ., data=mtcars)
summary(fit2)
```

```{r mtcars8}
cor(mtcars)[1,]
```

```{r mtcars9}
fit3 <- lm(mpg ~ wt+hp+disp+cyl+am, data=mtcars)
summary(fit3)
```

# Residual Analysis

```{r mtcars10}
par(mfrow = c(2,2))
plot(fit3)
```

Residual vs Fitted - no consistent pattern, supporting the accuracy of the independent assumption

Normal Q-Q plot - the residuals are normally distributed, because the points lie closely to the line.

Scale-Location - there is constant variance assumption, as the points are randomly distributed

Residuals vs Leverage - No outliners are present, as all values fall within the 0.5 bands

# Conclusion 

There is a difference in MPG based on the transmission type. A manual transmission will have a small MPG advantage than automatics ones. Weight, Horsepower, displacement, and number of cylinders are confounding variables in the relationship between transmission type and miles per gallon which manual transmission on average have 1.55 miles per gallon more than automatic cars.

